Towards a Pandemic-Scale Ancestral Recombination Graph of SARS-CoV-2

### 2023-04-24 paper.tex
Updates from Overleaf

\subsection{Imputation of Pango lineage for non-sample nodes}
Initially, only sample nodes have Pango lineage assignments.
For analysis of recombinant parents, it is also helpful to assign
Pango lineages to non-sample nodes. We do this
using a two-step procedure (Figure 3). If the Pango
lineage of the parental node was available and no mutation was immediately
ancestral to the non-sample node, then the parental node’s status was directly
propagated to the non-sample node. Otherwise, the Pango lineage of the
non-sample node was inferred by matching its full set of mutations against the
Pango lineage-defining mutations (based on 90\% consensus of the sequences
analysed) from the COVID-CG website \citep{Chen2021-zc};
\url{https://covidcg.org/}; accessed on November 04, 2022).

>>>>

\subsection{Imputation of Pango lineage for non-sample nodes}
Initially, only sample nodes have Pango lineage assignments.
For analysis of recombinant parents, it is also helpful to assign
Pango lineages to non-sample nodes. We do this
using a two-step procedure (Figure~\ref{fig:imputation}). If the Pango
lineage of the parental node was available and no mutation was immediately
ancestral to the non-sample node, then the parental node’s status was directly
propagated to the non-sample node. Otherwise, the Pango lineage of the
non-sample node was inferred by matching its full set of mutations against the
Pango lineage-defining mutations (based on 90\% consensus of the sequences
analysed) from the COVID-CG website \citep{Chen2021-zc};
\url{https://covidcg.org/}; accessed on November 04, 2022).

\begin{figure} \centering
\includegraphics[width=\textwidth]{figures/imputation.pdf}
\caption{\label{fig:imputation}A schematic of the two-step procedure to impute Pango lineage for inserted, non-sample nodes. Here, three nodes (question marks) have unknown Pango lineage (A). Because the parent of node 1 is available (i.e., L2) and there is no mutation occurring on the edge connecting the two nodes, the Pango lineage of the parent is directly propagated to node 1 (B). However, if there is a mutation (red X’s), then the Pango lineage is inferred by comparison to the set of lineage-defining mutations from COVID-CG\citep{Chen2021-zc}, as in case (C). Finally, the inferred status is propagated to any non-mutant child nodes, as in case (D).
}
\end{figure}
